Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending

### 2017-10-12 paper.Rmd
major update

---
title: Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending
author:
  - name: Joe Brew
    email: joebrew@economicsofmalaria.com
    affiliation: isglobal,cism
    footnote: Corresponding Author
  - name: Menno Pradhan
    email: m.p.pradhan@vu.nl
    affiliation: vu,uva
  - name: Jacqueline Broerse
    email: j.e.w.broerse@vu.nl
    affiliation: vu
  - name: Elisa Sicuri
    email: elisa.sicuri@isglobal.org 
    affiliation: isglobal,icl,cism
address:
  - code: isglobal
    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
  - code: icl
    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
  - code: cism
    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
  - code: vu
    address: Faculty of Economics and Business Administration, VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
  - code: uva
    address: Faculty of Economics and Business, Section Microeconomics, University of Amsterdam, REC E, Roetersstraat 11, Amsterdam, Netherlands
abstract: |
  Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a systematic survey of malaria researchers, we query perceptions regarding the likelihood and time frame of eradication, as well as the perceived chief areas needed for improvement in order for eradication to be achieved. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise. Our results, which suggest a disconnect between public institutional discourse and private opinion, serve as a barometer of professional perspectives, and serve to highlight areas where attention is most needed for eradication to be achieved.

bibliography: library.bib
params:
  get_new_data: False
output: rticles::elsevier_article
---

@@

# Get data
source('prepare_data.R')
```

@@

## Data generation

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers' demographics is of value. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name, and the likelihood of a participant being of one of five ethnicities (White, Black, Hispanic, Asian, Other) based on the provided last name.

@@

@@

Ordinary least squares regression was employed to estimate the effect of area of expertise on the eradication pessimism-optimism (through the proxy of perceived years to eradication).

Content analysis of free text comments was carried out using frequency analyses. Comments were categorized and coded according to (more details needed here).

All analysis was carried out in R language [@R]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey). The searching and retrieval of information from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

The estimation of gender and ethnicity likelihoods for each participant relied on Mullen's implementation of US Social Security Administration names' analysis [@gender2015] and Khanna et al's bayesian prediction algorithm for the estimation of racial categories [@wru2015].

De-biasing of sample selection was carried out using Heckman's 2 step approach as implemented by Toomet [@toomet2008].

@@

Respondents were qualitatively different from non-respondents. Importantly, the average number of total author-specific citations was `r round(mean(df$n_citations_total[df$responded]), digits = 2)` among respondents, but `r round(mean(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. This suggests either a tendency for more senior or more cited researchers to not respond. The distribution of author-specific total citations is shown below (a).

```{r}
g1 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_total)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6) +
  scale_x_log10() +
  theme_publication() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Total lifetime citations')
```

When examining the number of average citations per article, the difference between respondents was slightly less: `r round(median(df$n_citations_total[df$responded]), digits = 2)` among respondents, and `r round(median(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. Again, the distribution is shown below (b).

```{r}
g2 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_average)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6,
               n = 10) +
  scale_x_log10() +
  theme_publication() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Average citations per article')
Rmisc::multiplot(g1, g2, cols = 2)
```

The below table shows the response rate by different demographic groups of interest.

```{r}
# function to get response rate per categorical variable
response_rate <- function(variable = 'sex'){
  df$x <- df[,variable]
  out <- df %>%
    group_by(x) %>%
    summarise(responded = length(which(responded)),
              n = n()) %>%
    ungroup %>%
    mutate(p = responded / n * 100)
  return(out)
}
out_list <- list()
vars <- c('sex',
            'ethnicity',
            'bin_citations')
for (i in 1:length(vars)){
  this_var <- vars[i]
  x <- response_rate(variable = this_var)
  x$variable <- this_var
  out_list[[i]] <- x
}
out <- bind_rows(out_list) %>%
  dplyr::select(variable, 
                x,
                responded,
                n, 
                p) %>%
  rename(Variable = variable,
         Characteristic = x,
         Responded = responded,
         Invited = n,
         `Response rate` = p)
kable(out, format = 'latex', caption = 'Response rate by author characteristics')
```

The below chart shows response rate by number of citations.

```{r}
x <- out %>%
  filter(Variable == 'bin_citations') 
ggplot(data = x,
       aes(x = Characteristic,
           y = `Response rate`)) +
  geom_bar(stat = 'identity',
           fill = 'lightblue',
           alpha = 0.8) +
  theme_publication() +
  labs(x = 'Number of total citations',
       y = 'Response rate',
       title = 'Response rate by citations') +
  geom_label(aes(label = paste0(round(`Response rate`, digits = 2), '%')))
```

```{r, results = 'markup'}
# stargazer(ols1, heck1, ml1,
#           header = FALSE,
#           # se=list(cse(ols1),NULL,NULL),
#           title="Regression on years until eradication", 
#           type="text",
#           df=FALSE, digits=4)
# # summary(ols1)
# # summary(heck1)

@@

```{r}
# Define function for analyzing the years_n columns
years_n <- function(n){
  x <- df %>%
    filter(responded) %>%
    dplyr::select_(paste0('years_', n)) 
  names(x) <- 'likelihood'
  result <- x %>%
    group_by(likelihood) %>%
    tally %>%
    mutate(p = n / sum(n) * 100) %>%
    filter(!is.na(likelihood))

  out <- list()
  out[[1]] <- result
  out[[2]] <-
    ggplot(data = result,
           aes(x = likelihood,
               y = p)) +
    geom_bar(stat = 'identity',
             fill = 'darkorange',
             alpha = 0.6) +
    labs(title = paste0(toupper(letters[n/10]),'. ', n, ' years'),
         subtitle = 'Perceived likelihood of eradication',
         x = 'Likelihood',
         y = '%') +
    theme_publication(base_size = 10) +
    theme(axis.text.x = element_text(angle = 90))
  return(out)
}
```

```{r, fig.height = 7}
# Make plots of all
years <- seq(10, 50, 10)
for (i in 1:length(years)){
  assign(paste0('g',i ),
         years_n(n = years[i])[[2]])
  x <- years_n(n = years[i])[[1]]
  x$year <- years[i]
  assign(paste0('x', i),
         x)

@@

g6 <- 
  ggplot(data = xa,
         aes(x = year,
             y = avg)) +
  geom_ribbon(aes(ymin = q1,
                  ymax = q3),
              fill = 'orange',
              alpha = 0.6) +
  geom_line(color = 'darkorange') +
  theme_publication(base_size = 10) +
  xlab('Years') +
  ylab('Average likelihood') +
  ggtitle('G. Aggregated perceptions',
          'Shaded area = inter-quartile range')

@@

Participants were asked to provide the number of years they believe it will take until eradication can be achieved. `r length(which(is.na(df$years)))` (`r round(length(which(is.na(df$years))) / nrow(df) * 100, digits = 1)`%) were either blank or uninteligible, whereas `r length(which(!is.na(df$years)))` participants responded. Among respondants, `r length(which(df$years >= 50))` (`r round(length(which(df$years >= 50)) / length(which(!is.na(df$years))) * 100, digits = 1)` %) estimated that it would be 50 or more years until eradication. The below histogram shows the distribution of responses, appended at 50.

```{r, fig.height = 3.5, fig.width = 4}
ggplot(data = df,
       aes(x = years)) +
  geom_histogram(fill = 'darkorange',
                 alpha = 0.6) +
  theme_publication(base_size = 10) +
  labs(x = 'Years',
       y = 'Responses',
       title = 'Perceived years until eradication',
       subtitle = '(Estimates of greater than 50 years reduced to 50)')
```

@@

ggplot(data = dist_results,
       aes(#group = area,
           #fill = area,
           x = val)) +
  geom_density(fill = 'darkorange',
               alpha = 0.6) +
  theme_publication(base_size = 10) +
  labs(x = 'Years',
       y = 'Density',
       title = 'Perceived years to eradication',
       subtitle = '12 most popular areas of expertise only') +
  facet_wrap(~the_area) 
```

```{r, fig.height = 3.5, fig.width = 4}
model_formula <- paste0('years ~ ',
                        paste0(paste0('area_', gsub(' ', '_', small_areas)), collapse = ' + '))
model_formula <- as.formula(model_formula)
fit <- lm(model_formula, data = df)
# Get confidence intervals too
ci <- tidy(confint(fit))
x <- cbind(tidy(fit), ci)
names(x) <- c('term', 'estimate', 'se', 'stat', 'p', 'rownames', 'lwr', 'upr')
x <- x %>%
  dplyr::select(term, estimate,
                p, lwr, upr)
the_intercept <- x$estimate[1]
x <- x[-1,]
for (the_column in c('estimate', 'lwr', 'upr')){
  x[,the_column] <- x[,the_column] + the_intercept
}
x$term <- gsub('area_|TRUE', '', x$term)
x <- x %>%
  dplyr::arrange(estimate)
x$term <- gsub('_', '\n', x$term)
x$term <- factor(x$term, levels = x$term)
ggplot(data = x,
       aes(x = term,
           y = estimate)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  geom_errorbar(aes(ymin = lwr,
                    ymax = upr),
                alpha = 0.6,
                width = 0.8) +
  theme_publication(base_size = 10) +
  labs(x = 'Area',
       y = 'Years',
       title = 'Perceived years to eradication',
       subtitle = 'By expertise, adjusted for multiple areas of expertise') +
  theme(axis.text.x = element_text(angle = 90))

```{r, fig.height = 4, fig.width = 4.5}
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(year_results$area)))
ggplot(data = year_results,
       aes(x = year,
           y = val)) +
  geom_line(aes(group = area,
           color = area,
           lty = area)) +
  theme_publication(base_size = 10) +
  scale_color_manual(name = '',
                     values = cols,
                     labels = unique(year_results$area)) +
  scale_linetype_manual(name="", 
                        values = rep(1:6, 100)[1:length(unique(year_results$area))],
                     labels = unique(year_results$area)) +
  xlab('Years until eradication') +
  ylab('Likelihood') +
  ggtitle('Likelihood trajectories',
          'By area of expertise')
```

## Sentiment analysis of free-text commentary

This section is unfinished. Below is the distribution of "sentiment" among the responses (0 is neutral, below 0 is negative, above 0 is positive), but this is preliminary. The chart is colored by ethnicity, and faceted by gender.

```{r}
ggplot(data = df %>%
         filter(responded) %>%
         # mutate(n_areas = factor(n_areas)) %>%
         filter(!is.na(ethnicity),
                !is.na(gender)),
       aes(x = sentiment,
           group = ethnicity)) +
  geom_density(aes(fill = ethnicity),
               alpha = 0.6,
               n = 10) +
  theme_publication() +
  facet_wrap(~gender)

@@

ggplot(data = results) + 
  geom_violin(aes(x = bb,
           y = ranking),
           fill = 'orange',
              alpha = 0.6) +
  geom_jitter(aes(x = bb,
           y = ranking),
           color = 'black',
           with = 0.4,
           height = 0.4,
              alpha = 0.3,
              size = 1) +
  theme_publication(base_size = 10) +
  labs(x = 'WHO Building block',
       y = 'Ranking',
       title = 'Perceived areas needing attention',
       subtitle = 'Small points: jittered individual responses; large triangles: averages') +
  geom_point(data = results_agg,
             aes(x = bb,
                 y = avg),
             size = 5,
             pch = 17,
             color = 'black',
             alpha = 0.8) +
  ylim(0, 6)
ggsave('~/Desktop/figure2.png', height = 4, width = 5.5)
```

@@

Participants were invited to provide "any general comments on the timeline and likelihood of global eradication". Of the `r length(which(df$responded))` who responded to the survey, `r length(which(!is.na(df$comments)))` (`r round(length(which(!is.na(df$comments))) / length(which(df$responded)) * 100, digits = 2)`%) provided a comment.

@@

## Collective pessimism

@@

## Building blocks to eradication

@@

## Areas of expertise

@@

>>>>

---
title: Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending
bibliography: library.bib
output: 
  pdf_document:
    includes:
      in_header: preamble.sty
---

@@

# Get data
source('prepare_data.R', echo = FALSE)

theme_paper <- theme_publication
```

```{r}
authors <- data.frame(name = c('Joe Brew',
                               'Menno Pradhan',
                               'Jacqueline Broerse',
                               'Elisa Sicuri'),
                      email = c('joe@databrew.cc',
                                'm.p.pradhan@vu.nl',
                                'j.e.w.broerse@vu.nl',
                                'elisa.sicuri@isglobal.org'),
                      affilitation = c("isglobal,cism,vu",
                                       "vu,uva",
                                       "vu",
                                       "isglobal,icl,cism"),
                      footnote = c("Corresponding Author",
                                   rep('', 3)),
                      stringsAsFactors = FALSE)
institutions <- data.frame(code = c('isglobal',
                                    'icl',
                                    'cism',
                                    'vu',
                                    'uva'),
                           institution = c('Barcelona Institute for Global Health',
                                           'Imperial College London',
                                           'Centro de Investigação em Saúde de Manhiça',
                                           'VU University Amsterdam',
                                           'University of Amsterdam'),
                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
                                       'South Kensington Campus, London SW7 2AZ, U.K.',
                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands'),
                      stringsAsFactors = FALSE)

make_author <- function(authors,
                        institutions){

  out <- c()
  for (i in 1:nrow(authors)){
    this_author <- authors$name[i]
    these_institutions <- authors$affilitation[i]
    these_institutions <- unlist(strsplit(these_institutions, ','))
    author_text <- paste0(this_author)
    for (j in 1:length(these_institutions)){
      this_institution <- these_institutions[j]
      if(any(grepl(this_institution, out))){
        author_text <- paste0(author_text,
                              '\\footrecall{',
                              this_institution,
                              '}'
                              )
      } else {
        # new institution, get full name
        full_name <- institutions$institution[institutions$code == this_institution]
        address <- institutions$address[institutions$code == this_institution]
        author_text <- paste0(author_text,
                              '\\footremember{',
                              this_institution,
                              '}{',
                              full_name,
                              ': ',
                              address,
                              '}')
      }
    }
    out[i] <- author_text
  }
  out <- paste0(out, collapse = ', ')
  cat(out)
}
```

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_author(authors = authors,
            institutions = institutions)
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a systematic survey of malaria researchers, we query perceptions regarding the likelihood and time frame of eradication, as well as the perceived chief areas needed for improvement in order for eradication to be achieved. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise, and adjust for selection bias. Our results show a disconnect between optimistic public institutional discourse and pessimistic private opinion, suggesting either (a) the necessity of a realistic accounting of eradication's low short-term likelihood in health planning and financing, or (b) the need for a more compelling case regarding eradication's feasibility to be made to the research community.
\end{changemargin}
\end{center}

@@

## Supplementary data generation

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers is of value for the assessing of selection bias and determinants of perception. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name, and the likelihood of a participant being of one of five ethnicities (White, Black, Hispanic, Asian, Other) based on the provided last name. For gender estimations we used data from the North Atlantic Population Project, and U.S. government [@Lincoln]; for ethnicity, Khanna et al's bayesian prediction algorithm for the estimation of racial categories [@wru2015, @Imai_2016] was employed.

@@

The searching and retrieval of information from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

@@

Ordinary least squares regression was employed to estimate the effect of area of expertise on the eradication pessimism-optimism (through the proxy of perceived years to eradication). De-biasing of sample selection was carried out using Heckman's 2 step approach as implemented by Toomet [@toomet2008].

Qualitative analysis of free text comments was carried out using thematic content analysis.

Quantitative analysis was carried out in R language [@R]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

@@

@@

```{r, fig.cap='Perceptions of likelihood of eradication', fig.height = 6.2}
# Define function for analyzing the years_n columns
years_n <- function(n){
  x <- df %>%
    filter(responded) %>%
    dplyr::select_(paste0('years_', n)) 
  names(x) <- 'likelihood'
  result <- x %>%
    group_by(likelihood) %>%
    tally %>%
    mutate(p = n / sum(n) * 100) %>%
    filter(!is.na(likelihood))

  out <- list()
  out[[1]] <- result
  out[[2]] <-
    ggplot(data = result,
           aes(x = likelihood,
               y = p)) +
    geom_bar(stat = 'identity',
             fill = 'darkorange',
             alpha = 0.6) +
    labs(title = paste0(toupper(letters[n/10]),'. ', n, ' years'),
         subtitle = 'Perceived likelihood of eradication',
         x = 'Likelihood',
         y = '%') +
    theme_paper(base_size = 10) +
    theme(axis.text.x = element_text(angle = 90))
  return(out)
}

# Make plots of all
years <- seq(10, 50, 10)
for (i in 1:length(years)){
  assign(paste0('g',i ),
         years_n(n = years[i])[[2]])
  x <- years_n(n = years[i])[[1]]
  x$year <- years[i]
  assign(paste0('x', i),
         x)

@@

g6 <- 
  ggplot(data = xa,
         aes(x = year,
             y = avg)) +
  geom_ribbon(aes(ymin = q1,
                  ymax = q3),
              fill = 'orange',
              alpha = 0.6) +
  geom_line(color = 'darkorange') +
  theme_paper(base_size = 10) +
  xlab('Years') +
  ylab('Average likelihood') +
  ggtitle('G. Aggregated perceptions',
          'Shaded area = inter-quartile range')

@@

Participants were asked to provide the number of years they believe it will take until eradication can be achieved. `r length(which(is.na(df$years[df$responded])))` (`r round(length(which(is.na(df$years[df$responded]))) / nrow(df) * 100, digits = 1)`%) were either blank or uninteligible, whereas `r length(which(!is.na(df$years)))` participants responded. Among respondants, `r length(which(df$years >= 50))` (`r round(length(which(df$years >= 50)) / length(which(!is.na(df$years))) * 100, digits = 1)` %) estimated that it would be 50 or more years until eradication. The below histogram shows the distribution of responses, truncated at 50.

```{r, fig.height = 3.5, fig.width = 4}
ggplot(data = df,
       aes(x = years)) +
  geom_histogram(fill = 'darkorange',
                 alpha = 0.6) +
  theme_paper(base_size = 10) +
  labs(x = 'Years',
       y = 'Responses',
       title = 'Perceived years until eradication',
       subtitle = '(Estimates of greater than 50 years reduced to 50)')
```

Respondents were qualitatively different from non-respondents. Importantly, the average number of total author-specific citations was `r round(mean(df$n_citations_total[df$responded]), digits = 2)` among respondents, but `r round(mean(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. This suggests either a tendency for more senior or more cited researchers not to respond. When examining the number of average citations per article, the difference between respondents and non-respondents was slightly less: `r round(median(df$n_citations_total[df$responded]), digits = 2)` among respondents, and `r round(median(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. Males responded at a greater rate (`r round(mean(df$responded[df$sex == 'male'], na.rm = T) * 100, digits = 2)`) than females (`r round(mean(df$responded[df$sex == 'female'], na.rm = T) * 100, digits = 2)`) and those with Hispanic last names responded at a greater rate than those with last names of other ethnicities (see table).

```{r}
# function to get response rate per categorical variable
response_rate <- function(variable = 'sex'){
  df$x <- df[,variable]
  out <- df %>%
    group_by(x) %>%
    summarise(responded = length(which(responded)),
              n = n()) %>%
    ungroup %>%
    mutate(p = responded / n * 100)
  return(out)
}
out_list <- list()
vars <- c('sex',
            'ethnicity',
            'bin_citations')
for (i in 1:length(vars)){
  this_var <- vars[i]
  x <- response_rate(variable = this_var)
  x$variable <- this_var
  out_list[[i]] <- x
}
out <- bind_rows(out_list) %>%
  dplyr::select(variable, 
                x,
                responded,
                n, 
                p) %>%
  rename(Variable = variable,
         Characteristic = x,
         Responded = responded,
         Invited = n,
         `Response rate` = p)
knitr::kable(out, format = 'latex', caption = 'Response rate by author characteristics')
```

Selection bias is not of concern in the case of differential response if the groups for whom there are differences are not different in terms of the outcome variable. This was the case for sex: males responded at a significantly greater rate than females (p < 0.001), but were not statistically significantly different in regards to pessimism/optimism (ie, time frame or likelihood of eradication). In the case of researcher impact (as measured by the total number of citations), selection bias plays an important role: being highly-cited was associated both with eradication "pessimism" as well as likelihood of non-response. In other words, our pool of respondents was less highly-cited than our pool of invitees, and among respondents, those who were highly-cited tended to be more pessimistic (see below figure).

```{r}
# df$bin_citations <- factor(df$bin_citations,
#                            labels = c('0',
#                                       '1-9',
#                                       '10-51',
#                                       '52+'))
x <- df %>%
  group_by(citations = bin_citations) %>%
  summarise(years_10 = mean(years_10_numeric, na.rm = TRUE),
            years_20 = mean(years_20_numeric, na.rm = TRUE),
            years_30 = mean(years_30_numeric, na.rm = TRUE),
            years_40 = mean(years_40_numeric, na.rm = TRUE),
            years_50 = mean(years_50_numeric, na.rm = TRUE),
            n = length(unique(email))) %>%
  ungroup %>%
  gather(key, value, years_10:years_50) %>%
  filter(!is.na(citations))

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = '% likelihood',
       title = 'Difference in perceived likelihood of eradication by citations',
       subtitle = 'More citations = more pessimistic')
```

We used Heckman's 2-step correction method [@Heckman_1979]. Given that sex affected response rate but not response itself, we were able to use sex as an instrumental variable, introducing it in both the selection and response models. Our selection model, which estimated the likelihood of survey response as a probit function of sex allowed us to estimate (via the Inverse Mills ratio) an expected error term for our outcome models, in which we quantify the effect of the number of citations on the perceived average likelihood of eradication at 10, 20, 30, 40, and 50 years. The below chart shows the difference in the aggregate perceived likelihood of eradication before and after implementing Heckman's 2-step model.

```{r, results = 'markup'}
ggplot(data = heckman,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
  geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange')) +
  labs(title = 'Heckman selection adjustment')
```

The below table shows the Heckman model Heckman model estimates for all five models.

```{r}
rn <- names(heck_10$coefficients)
heck <- data_frame(Coefficient = rn,
                   `Model 10` = heck_10$coefficients,
                   `Model 20` = heck_20$coefficients,
                   `Model 30` = heck_30$coefficients,
                   `Model 40` = heck_40$coefficients,
                   `Model 50` = heck_50$coefficients)
knitr::kable(heck)
```

@@

```{r, fig.height = 3.5, fig.width = 4}
areas_df <- areas_df %>% arrange(desc(count))
small_areas <- areas_df$area[1:12]
model_formula <- paste0('years ~ ',
                        paste0(paste0('area_', gsub(' ', '_', small_areas)), collapse = ' + '))
model_formula <- as.formula(model_formula)
fit <- lm(model_formula, data = df)
# Get confidence intervals too
ci <- tidy(confint(fit))
x <- cbind(tidy(fit), ci)
names(x) <- c('term', 'estimate', 'se', 'stat', 'p', 'rownames', 'lwr', 'upr')
x <- x %>%
  dplyr::select(term, estimate,
                p, lwr, upr)
the_intercept <- x$estimate[1]
x <- x[-1,]
for (the_column in c('estimate', 'lwr', 'upr')){
  x[,the_column] <- x[,the_column] + the_intercept
}
x$term <- gsub('area_|TRUE', '', x$term)
x <- x %>%
  dplyr::arrange(estimate)
x$term <- gsub('_', '\n', x$term)
x$term <- factor(x$term, levels = x$term)
ggplot(data = x,
       aes(x = term,
           y = estimate)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  geom_errorbar(aes(ymin = lwr,
                    ymax = upr),
                alpha = 0.6,
                width = 0.8,
                color = 'red') +
  theme_paper(base_size = 10) +
  labs(x = 'Area',
       y = 'Years',
       title = 'Perceived years to eradication',
       subtitle = 'By expertise, adjusted for multiple areas of expertise') +
  theme(axis.text.x = element_text(angle = 90))

@@

library(ggridges)
ggplot(data = results %>% filter(!is.na(ranking)),
       aes(x = ranking,
           y = bb,
           fill = bb)) +
  geom_density_ridges(#fill = 'darkorange',
                      alpha = 0.6,
                      scale = 1.5) +
  theme_paper() +
  xlim(0,30) +
  labs(x = 'Perceived importance',
       y = 'WHO Building Block') +
 scale_fill_cyclical(values = c("darkorange", "yellow"))

# ggplot(data = results) + 
#   geom_violin(aes(x = bb,
#            y = ranking),
#            fill = 'orange',
#               alpha = 0.6) +
#   geom_jitter(aes(x = bb,
#            y = ranking),
#            color = 'black',
#            with = 0.4,
#            height = 0.4,
#               alpha = 0.3,
#               size = 1) +
#   theme_paper(base_size = 10) +
#   labs(x = 'WHO Building block',
#        y = 'Ranking',
#        title = 'Perceived areas needing attention',
#        subtitle = 'Small points: jittered individual responses; large triangles: averages') +
#   geom_point(data = results_agg,
#              aes(x = bb,
#                  y = avg),
#              size = 5,
#              pch = 17,
#              color = 'black',
#              alpha = 0.8) +
#   ylim(0, 6)
```

@@

Participants were invited to provide "any general comments on the timeline and likelihood of global eradication". Of the `r length(which(df$responded))` who responded to the survey, `r length(which(!is.na(df$comments)))` (`r round(length(which(!is.na(df$comments))) / length(which(df$responded)) * 100, digits = 2)`%) provided a comment. Thematic content analysis [@Vaismoradi_2016] was employed to code responses into four subject themes as well as binary categorization of descriptive vs. presciptive.

The four subject themes, identified and refined through iterative analysis, were:

1. Innovation
2. Systematic challenges  
3. Poverty
4. Meta: challenging the question, re-framing the goal, insisting on other priorities, etc.

@@

## General pessimism

@@

## Discord over building blocks to eradication

@@

## Few differences by area of expertise

@@

# Appendix

The distribution of author-specific total citations is shown below (a).

```{r}
g1 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_total)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6) +
  scale_x_log10() +
  theme_paper() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Total lifetime citations')
```

Again, the distribution is shown below (b).

```{r}
g2 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_average)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6,
               n = 10) +
  scale_x_log10() +
  theme_paper() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Average citations per article')
# Rmisc::multiplot(g1, g2, cols = 2)
g1
g2
```

ggplot(data = dist_results,
       aes(#group = area,
           #fill = area,
           x = val)) +
  geom_density(fill = 'darkorange',
               alpha = 0.6) +
  theme_paper(base_size = 10) +
  labs(x = 'Years',
       y = 'Density',
       title = 'Perceived years to eradication',
       subtitle = '12 most popular areas of expertise only') +
  facet_wrap(~the_area) 
```

```{r, fig.height = 4, fig.width = 4.5}
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(year_results$area)))
ggplot(data = year_results,
       aes(x = year,
           y = val)) +
  geom_line(aes(group = area,
           color = area,
           lty = area)) +
  theme_paper(base_size = 10) +
  scale_color_manual(name = '',
                     values = cols,
                     labels = unique(year_results$area)) +
  scale_linetype_manual(name="", 
                        values = rep(1:6, 100)[1:length(unique(year_results$area))],
                     labels = unique(year_results$area)) +
  xlab('Years until eradication') +
  ylab('Likelihood') +
  ggtitle('Likelihood trajectories',
          'By area of expertise')
```

```{r}
ggplot(data = df %>%
         filter(responded) %>%
         # mutate(n_areas = factor(n_areas)) %>%
         filter(!is.na(ethnicity),
                !is.na(gender)),
       aes(x = sentiment,
           group = ethnicity)) +
  geom_density(aes(fill = ethnicity),
               alpha = 0.6,
               n = 10) +
  theme_paper() +
  facet_wrap(~gender)
